\section{Introduction:}\label{sec:introduction}
% Singapore, a frontrunner in sustainable urban development, grapples with a crucial data gap: the lack of a dedicated Wireless Sensor Network (WSN) to monitor harmful gas (CO\textsubscript{2}) emissions and their intricate link to temperature fluctuations across diverse urban environments. This absence of comprehensive data impedes our ability to accurately track progress towards ambitious environmental targets and formulate informed policies for critical issues like CO\textsubscript{2} reduction and urban heat island mitigation.
The absence of a dedicated Wireless Sensor Network (WSN) for monitoring harmful gas emissions, including (CO\textsubscript{2}), and their relation to temperature fluctuations in diverse urban environments is a critical gap that has been recognized in various contexts, not specifically limited to Singapore. However, studies like "UrbanSense" \cite{Rainham2016A} and "CitySee" \cite{Liu2013CitySee} have demonstrated the urgent need for and the implementation of WSN infrastructures in urban areas to monitor environmental conditions, including (CO\textsubscript{2}) and temperature fluctuations, which are crucial for formulating informed policies towards (CO\textsubscript{2}) reduction and urban heat island mitigation.

\section{Problem Statement}\label{sec:problem_statement}
The lack of comprehensive environmental monitoring poses a significant challenge within Singapore. Currently, there is a dearth of real-time data on CO\textsubscript{2} levels and temperature, hindering informed decision-making and sustainable practices in the country. This data gap inhibits the establishment of correlations between environmental factors and their impact on campus well-being and sustainability. Consequently, there is a critical need to develop and deploy a scalable and energy-efficient mesh network utilizing efficient communication protocols to address this challenge. 

\subsection{Objectives}\label{sec:objectives}

This project aims to develop a robust multi-protocol mesh system for IoT applications, with a focus on real-time environmental data monitoring. The objectives are as follows:

\begin{enumerate}
    \item Design and implement a flexible IoT architecture capable of seamlessly integrating multiple protocols, including LoRa, ESP-Now, and traditional mesh algorithms.
    \item Develop adaptive routing algorithms to dynamically switch between protocols based on environmental conditions, network congestion, and distance variations, ensuring continuous data transmission and network reliability.
    \item Evaluate the performance, scalability, and reliability of each protocol in real-world scenarios, considering factors such as data latency, packet loss, and network reach.
    \item Implement a centralized dashboard for visualizing and analyzing environmental data collected from sensor nodes, providing stakeholders with actionable insights for informed decision-making.
    \item Integrate mechanisms for automatic protocol fallback and toggling in response to network failures, distance changes, or protocol-specific limitations, ensuring uninterrupted data flow and system resilience.
    \item Conduct extensive testing and validation of the multi-protocol mesh system in a simulated and real-world environment, focusing on its ability to adapt to dynamic network conditions and maintain data integrity.
    % \item Collaborate with policymakers, stakeholders, and environmental experts to translate collected data into evidence-based strategies for enhancing environmental sustainability within NYP Campus and supporting broader sustainability initiatives in Singapore.
\end{enumerate}

By achieving these objectives, the project aims to establish a scalable and reliable IoT infrastructure capable of supporting diverse applications while promoting data-driven decision-making and sustainable management practices in alignment with Singapore's environmental goals.

% \subsection{Objectives}\label{sec:objectives}

% By providing real-time environmental data, this project aims to enable:

% \begin{itemize}
%     \item Identification of correlations between CO\textsubscript{2} levels, temperature, and campus environmental conditions.
%     \item Evaluation of the performance and suitability of LoRa and ESP-Now protocols in comparison to traditional mesh algorithms, addressing concerns regarding data latency, reliability, network scalability, and reach.
%     \item Provision of actionable insights to policymakers and stakeholders, facilitating evidence-based strategies for enhancing environmental sustainability within NYP Campus and contributing to broader sustainability initiatives in Singapore.
% \end{itemize}

% Through the resolution of this pressing problem, the project endeavors to pave the way for data-driven decision-making and sustainable management practices within NYP Campus, aligning with Singapore's commitment to environmental stewardship.

% \section*{Problem Statement}\label{sec:problem_statement}
% This project tackles this pressing challenge by proposing the development and deployment of a scalable and energy-efficient mesh network utilizing LoRa and ESP-Now protocols in Nanyang Polytechnic (NYP) Campus. Through this network, we aim to collect and analyze real-time CO\textsubscript{2} and temperature data, enabling us to achieve three key objectives:
% 
% \begin{itemize}
%     \item Establish robust correlations between these environmental factors,
%     \item Evaluate the performance and suitability of LoRa and ESP-Now protocols compared to established mesh algorithms, and
%     \item Deliver valuable insights to policymakers and stakeholders, empowering them to develop data-driven strategies for a more sustainable urban future.
% \end{itemize}
% 
% By addressing this data gap and providing actionable insights, this project aspires to contribute significantly to Singapore's journey towards environmental sustainability.
% 
% \subsection*{Focus Areas}\label{sec:focus_areas}
% We will focus on key factors like:
% \begin{itemize}
%     \item \textbf{Data latency and reliability:} Ensure timely and accurate transmission of environmental data.
%     \item \textbf{Network scalability and reach:} Ability to handle a large number of nodes and cover the desired area effectively.
%     \item \textbf{Energy efficiency:} Minimize power consumption of sensor nodes for extended lifespan and network sustainability.
%     \item \textbf{Cost-effectiveness:} Consider hardware, deployment, and operational costs for a sustainable solution.
% \end{itemize}
